---
title: 'Activity: Popular Names'
author: "Haichen Wei"
date: "Oct 8, 2020"
output:
  html_document:
    df_print: paged
---



# Set Up

```{r}
# load necessary packages
library(tidyverse)
library(mosaic)
library(DataComputing)
```


```{r}
# Load the BabyNames data into RStudio environment (from DataComputing package)
data("BabyNames")

```

# Activity

**Goal**: By the end of the activity, each of you will create a graph (like Figure A.1 on p 159) showing ups and downs in the popularity of four names (your name & three additional names) according to the U.S. social security administration  


### Step 1

```{r}
summary(BabyNames)
```

### Step 3

• Sex does not appear in the graph.

• The count variable is transformed to popularity.

### Step 4

• The cases will be filtered out by specific names.

• Then, group the name and year, and summarized the sum of count within groups.

• The new variable is popularity which calculates the total count of each name within a year.

### Step 5

Create a GlyphReadyForm data frame, and choose four names by using filter function, then group the name and year, summarise the sum of name count. Finally, use ggplot() command to make graphic.

### Step 7

```{r}
GlyphReadyForm <-
  BabyNames %>%
    filter(name %in% c("Mary", "Anna", "Emma", "Haichen")) %>%
    group_by(name, year) %>%
    summarise(total= sum(count))
GlyphReadyForm %>% 
  ggplot(aes(x = year, y = total, group = name)) +
  geom_line( size = 1, alpha = 0.5, aes(color = name)) +
  ylab("Popularity") + xlab("Year")
```
