---
title: "STAT 184 Homework Exercises"
author: "Haichen Wei"
subtitle: DataComputing Chapters 11
output:
  html_document:
    df_print: paged
---

## Assigned Exercises

- DC Chapter 11 Exercises: 11.1, 11.2, 11.6

## Front Matter

```{r}
# clean up workspace environment
rm(list = ls())

# all packages used for the assignment
library(mosaic)
library(tidyverse)

```



## Chapter 11 Exercises

### Problem 11.1

Previous data verbs take as input a single data frame along with other arguments to specify the exact operation to be carried out. Join family data verbs take two tables as input and produce a single table as output.

### Problem 11.2

They have different numbers of rows, and the second row in Demographics is a different case entirely from the second row in CountryCentroids.

### Problem 11.6

```{r}
library(DataComputing)
data("ZipGeography")
help("ZipGeography")
view(ZipGeography)
```

1. 1909 counties
```{r}
ZipGeography %>%
  summarise(total = n_distinct(County, na.rm = TRUE))
```

2. Franklin is used in the most states.
```{r}
ZipGeography %>%
  group_by(CityName) %>%
  summarise(total = n_distinct(State, na.rm = TRUE)) %>%
  arrange(desc(total)) %>%
  head(2)
```

3. 
```{r}
ZipGeography %>%
  select(State, CityName, Population)%>%
  filter(Population >0 & State != "")%>%
  group_by(State)%>%
  summarise(totalS= sum(Population))%>%
  left_join(
    ZipGeography %>%
    select(State, CityName, Population)%>%
      filter(Population >0 & State != "")%>%
    group_by(State,CityName)%>%
      summarise(totalC=sum(Population))
  )%>%
  mutate(ratio = totalC/totalS)%>%
  arrange(desc(ratio))%>%
  filter(ratio>0.05)%>%
    group_by(CityName)%>%
    summarise(count = n_distinct(ratio))%>%
    arrange(desc(count))
```
4. 
```{r}
ZipGeography%>%
  select(State, Timezone)%>%
  group_by(State)%>%
  summarise(count = n_distinct(Timezone))%>%
  filter(count>1)
```
5. 
```{r}
ZipGeography%>%
  select(CityName, Timezone)%>%
  group_by(CityName)%>%
  summarise(count = n_distinct(Timezone))%>%
  filter(count>1)
```
6. 
```{r}
ZipGeography%>%
  select(County, Timezone)%>%
  group_by(County)%>%
  summarise(count = n_distinct(Timezone))%>%
  filter(count>1)
```
