---
title: "STAT 184 Homework Exercises"
author: "Haichen Wei"
subtitle: DataComputing Chapters 12
output:
  html_document:
    df_print: paged
---

## Assigned Exercises

- DC Chapter 12 Exercises: 12.1 (1, 2, 3), 12.2 (1, 2, 3, 4), 12.3 (1, 2), 12.4

## Front Matter

```{r}
# clean up workspace environment
rm(list = ls())

# all packages used for the assignment
library(mosaic)
library(tidyverse)
library(DataComputing)
```



## Chapter 12 Exercises

### Problem 12.1

1. sex
```{r}
BothSexes <-
  BabyNames %>%
  spread( key=sex, value=count ) %>%
  filter( F>1, M>1 )
```
2. 
```{r}
BothSexes%>%
  group_by(name)%>%
  summarise(F = sum(F), M = sum(M)) %>%
  mutate(gender_balance  = abs(log(F/M)) ) %>%
  arrange(gender_balance ) %>%
  head(10)
```
3. 
```{r}
BothSexes%>%
  mutate(gender_balance  = abs(log(F/M)) ) %>%
  arrange(gender_balance ) %>%
  filter(M > 100, F > 100) %>%
  head(10)
```

### Problem 12.2

1. • Version One: The number of baby names

• Version Two:Each sex's number of name

• Version Three: For the year 1912 and 2012, count the number of each name 

2. Version Two is wider.
```{r}
Version_Two<-
  BabyNames %>%
  spread(key = sex, value=count ) %>%
  filter(year %in% c(1912, 2012) )
Version_Two
```
3. Version Three is wider.
```{r}
Version_Three<-
  BabyNames %>%
  filter(year %in% c(1912, 2012) )  %>%
  spread(key = year, value=count)
Version_Three
```

```{r}
Version_One<-
  Version_Three %>%
    pivot_longer(cols = c("1912", "2012"), names_to = "year", values_to = "count")
Version_One
```

4. Version Two, because it has a single column for count of female and male separately, then we can directly use pmin(M / F, F / M) to calculate the ratio.

### Problem 12.3

1. • A is wider.

• B is wider.

• A and B are both wide format for the C table, so they can't be wider for each other.

2. Data Frame B is easiest to find the change. We can use mutate(change = Y2001 - Y2000) to add a new column. 

### Problem 12.4

1. The narrow data frame is difficult to find out the useful information at the first point, the information is messy and hard to read. 

2. BP_wide %>% gather(key = when, value = sbp, before, after )

