---
title: "STAT 184 Homework Exercises" 
subtitle: "DataComputing Chapters 7"
author: "Haichen Wei"
output: html_notebook
---

## Assigned Exercises

- DC Chapter 7 Exercises: 7.1, 7.2, 7.3, 7.5, 7.6  

## Front Matter

```{r}
# clean up workspace environment
rm(list = ls())

# all packages used for the assignment
library(mosaic)
library(tidyverse)

```



## Chapter 7 Exercises

### Problem 7.1

1. reduction function
2. neither
3. transformation function
4. transformation function
5. transformation function
6. reduction function
7. transformation function
8. transformation function

### Problem 7.2

a. First should not be in quotations. 
BabyNames %>% group_by(First) %>% summarise( votesReceived=n() )
b. It should refer to dataset before group by, and n() is the function used for count. Tmp <- BabyNames %>% group_by(BabyNames, year, sex ) %>% summarise( Tmp, totalBirths=n())
c. Same problems as part b, and missing the connection %>%.
Tmp <- BabyNames %>% group_by(BabyNames, year, sex) %>% summarise( BabyNames, totalBirths=n() )

### Problem 7.3
```{r}
library(DataComputing)
help(Minneapolis2013)
```
1. 80101 cases
2.
```{r}
Minneapolis2013 %>%
  group_by(Second) %>%
  tally(sort=TRUE)
```
   1 BETSY HODGES       14399
   2 DON SAMUELS        14170
   3 MARK ANDREW        12757
   4 undervote          10598
   5 JACKIE CHERRYHOMES  6470
3. First: 834
```{r}
Minneapolis2013 %>% 
  group_by(First) %>%
  filter(First=="undervote") %>%
  nrow()
```
   Second: 10598
```{r}
Minneapolis2013 %>% 
  group_by(Second) %>%
  filter(Second=="undervote") %>%
  nrow()
```
   Third: 19210
```{r}
Minneapolis2013 %>% 
  group_by(Third) %>%
  filter(Third=="undervote") %>%
  nrow()
```
4. 
```{r}
Minneapolis2013 %>%
  group_by(First,Second) %>% 
  summarise(count = n()) %>%
  arrange(desc(count))
```
BETSY HODGES	MARK ANDREW	8787
BETSY HODGES	DON SAMUELS	8743
MARK ANDREW	BETSY HODGES	6970
5. 
```{r}
Minneapolis2013 %>%
  group_by(Precinct) %>% 
  summarise(count = n()) %>%
  arrange(desc(count))
```
The precinct with the highest number of ballots is P-06 with 9711 ballots.

### Problem 7.5

a. Voterdata_A shows the combinations of First and Second vote selections.
b. VoterData_B shows the number of total cases.
c. VoterData_C shows the combinations of First and Second vote selections, and the total of combinations.
d. In VoterData_D, the summarise function has calculated the count of total cases, so it regards the entire dataset as group, it can't break down a subgroup when it's still grouped.

### Problem 7.6

```{r}
Pop <- ZipGeography[!is.na(ZipGeography$Population),]
tapply(Pop$Population, Pop$State, sum)

Area <- ZipGeography[!is.na(ZipGeography$LandArea),]
tapply(Area$LandArea, Area$State, sum)

```