---
title: "STAT 184 Homework Exercises"
author: "Haichen Wei"
subtitle: DataComputing Chapters 8
output:
  html_document:
    df_print: paged
---

## Assigned Exercises

- DC Chapter 8 Exercises: 8.1, 8.2, 8.3

## Front Matter

```{r}
# clean up workspace environment
rm(list = ls())

# all packages used for the assignment
library(mosaic)
library(tidyverse)

```



## Chapter 8 Exercises

### Problem 8.1

1. geom_point(): Add a layer of glyphs
2. geom_histogram(): Add a layer of glyphs
3. ggplot(): Construct a graphics frame
4. scale_y_log10(): Change the scale for the frame
5. ylab(): Set an axis label
6. facet_wrap(): Divide the frame into facets
7. geom_segment(): Add a layer of glyphs
8. xlim(): Set an axis label
9. facet_grid(): Divide the frame into facets

### Problem 8.2

```{r}
library(mosaicData)
library(DataComputing)
```

(A)
```{r}
ggplot( data = CPS85, aes(x = wage)) + geom_density(adjust = 1,alpha=0.5) + aes(colour = sex,fill=sex) + facet_wrap(~married, ncol = 4) + labs(title = "") + theme(legend.position = "right") 
```

(B)
```{r}
ggplot(data=CPS85,aes(x=age,y=wage))+aes(colour=sex)+facet_grid(married ~ .) + stat_smooth(method=loess) + ylim(0,15) 
```

(C)
```{r}
ggplot( data = CPS85, aes(x = age,y=wage)) + aes(colour = sex) + facet_wrap(~sector, ncol = 4) + stat_smooth(method=loess) + ylim(0,25) 
```

### Problem 8.3

(A)
```{r}
ggplot(data=CPS85,aes(x=age,y=wage))+geom_point() 
```

(B)
```{r}
ggplot(data=CPS85,aes(x=age,y=wage))+geom_point(aes(shape=sex)) 
```

(C)
```{r}
ggplot(data=CPS85,aes(x=age,y=wage))+geom_point(aes(shape=sex))+facet_grid(married ~ .) 
```

(D)
```{r}
ggplot(data=CPS85,aes(x=age,y=wage))+geom_point(aes(shape=married)) 
```
