---
title: "STAT 184 Homework Exercises"
author: "Haichen Wei"
subtitle: DataComputing Chapters 5 & 6
output:
  html_document:
    df_print: paged
---

## Assigned Exercises

- DC Chapter 5 Exercises: 5.1 
- DC Chapter 6 Exercises: 6.1 through 6.10

## Front Matter

```{r}
# clean up workspace environment
rm(list = ls())

# all packages used for the assignment
library(mosaic)
library(tidyverse)

```



## Chapter 5 Exercises

### Problem 5.1

```{r}
library(mosaicData)
library(DataComputing)
barGraphHelper(CPS85)

ggplot(data=CPS85,aes(x=sex,y=wage,fill=married))+
  geom_bar(stat='identity',position='dodge', width=.9)+ facet_wrap(~race,ncol=3)
```
## Chapter 6 Exercises

### Problem 6.1

1. suspension growth and adhesion growth
2. Yes, the frames are the same
3. red line:suspension growth for different amounts of time
   black lines: adhesion growth
   starts

### Problem 6.2

1. log enzyme concentration: y axis, range from -3 to 5
   flux:a filled dot or a hollow dot
   target: lines on top of the graph indicating the region in which each level
   gene: x axis.
   molecule: the different colors of the dot glyphs.
2. Color: different types of molecules 
   Fill: describes whether the molecule has flux = 0 or flux > 0. 
   Location: x-axis (gene type) and y-axis (log enzyme concentration)
3. Log enzyme concentration and gene
4. target and gene

### Problem 6.3

1. vertical lines: center,low,high
   star: polarity
   color: different types of protein
2. protein and cell density
3. Yes, color shows the dtfferent types of protein
4. Log2 ratio low vs high cell density, Apico-basal cell polarity proteins, color

### Problem 6.4

1. It's a horizontal coordinate which uses time and modes of transportation to indicate the position.
2. It uses three coclors(red, yellow,blue) to indicate different modes of transportation.
3. The shape/style of the glyphs encodes modes of transportation.
   The position of the glyph encodes the time.
4. time and modes of transportation, color

### Problem 6.5

b

### Problem 6.6

d

### Problem 6.7

b

### Problem 6.8

a, e

### Problem 6.9

```{r}
Small <- 
  NCHS %>% 
  sample_n(size=5000) 
  scatterGraphHelper(Small)
  ggplot(data=Small,aes(x=bmi,y=weight))+geom_point()+aes(colour=smoker)
```

Variables: bmi, weight, smoker
Attributes: bmi = x-axis, weight = y-axis, smoker = color

### Problem 6.10

a. The layer shows all the organisms, and some of them are overlapping
b. There is no scale in this graph.
