---
title: "STAT 184 Homework Exercises"
author: "Haichen Wei"
subtitle: DataComputing Chapters 15
output:
  html_document:
    df_print: paged
---

## Assigned Exercises

- DC Chapter 15 Exercises: 15.1, 15.2, 15.3, 15.4, 15.5  

## Front Matter

```{r}
# clean up workspace environment
rm(list = ls())

# all packages used for the assignment
library(mosaic)
library(tidyverse)
library(DataComputing)
```



## Chapter 15 Exercises

### Problem 15.1

```{r}
CPS85 %>%
  ggplot(aes(x=sex, y=wage)) + aes(fill=sex) + geom_boxplot()
```

### Problem 15.2

1. The most likely height will be 1.61 meters because the peak of the graph is there.
2. The most likely height for women is 1.61 m, for men is 1.64 m.
3. The most likely BMI for a 40-year old is 26, for a 70-year old is 27.

### Problem 15.3

```{r}
CPS85 %>%
  ggplot(aes(x=educ, y=wage)) + geom_point() + aes(colour=sex) + stat_smooth(method=lm) + ylim(0,15)
```

### Problem 15.4

```{r}
Galton %>%
  ggplot(aes(x=height, y=mother)) + geom_density2d() +facet_wrap(~sex, ncol = 4)
```

### Problem 15.5

• study 20 ZIP codes

1. No, the scattered plot is nonuniform.
2. Yes, the line is almost parallel to the x-axis.

• study 80 ZIP codes

1. The flat line is not consistent with the data. 
2. The function for 80 zip codes has a narrow confidence band, and 20 zip codes confidence band is wider. We can estimate it as 21/9, so the ratio of confidence band heights is about 2.3.
3. (1/√20)/ (1/√80) = 2, so this statistical theory seems right.

