---
title: "STAT 184 Homework Exercises"
author: "Haichen Wei"
subtitle: DataComputing Chapters 17
output:
  html_document:
    df_print: paged
---

## Assigned Exercises

- DC Chapter 17 Exercises: 17.1, 17.2, 17.3, 17.4, 17.5, 17.8

## Front Matter

```{r}
# clean up workspace environment
rm(list = ls())

# all packages used for the assignment
library(mosaic)
library(tidyverse)
library(DataComputing)
```



## Chapter 17 Exercises

### Problem 17.1

• Boys’ names ending in a vowel.
```{r}
BabyNames %>%
  filter(sex=="M", grepl( "[aeiou]$", name, ignore.case = TRUE))%>%
  group_by(name) %>%
  summarise(total = sum(count)) %>%
  arrange(desc(total)) %>%
  head(10)
```

• Names with a space (like Lee Ann)
```{r}
BabyNames %>%
  filter(grepl("\\s+", name, ignore.case = TRUE)) %>%
  group_by(name) %>%
  summarise(total = sum(count)) %>%
  arrange(desc(total)) %>%
  head(10)
```

• Names ending with “joe” (like BettyJoe)
```{r}
BabyNames %>%
  filter(grepl("joe$", name, ignore.case = TRUE)) %>%
  group_by(name) %>%
  summarise(total = sum(count)) %>%
  arrange(desc(total)) %>%
  head(10)
```

### Problem 17.2

```{r}
regex <- "^K[aeiou]+tl[aeiou]+n$|^C[aeiou]+tl[aeiou]+n$"
```

### Problem 17.3

It's like phone number, for example: 923-456-1234 or 923.456.1234 or 923 456 1234. It extracts those three pieces from strings that match in the parentheses, so the whole expression will be "one digit 2 to 9 followed by two digits 0 to 9, dash/space/period, three digits 0 to 9, dash/space/period, four digits 0 to 9".

### Problem 17.4

1. 2
2. 7+3+2+2+2+1+4+2+1+10+8+3+3+1+2+2+1+3+4=61 strings
3. The abbreviation of US states. 

### Problem 17.5

```{r}
BibleNames <- 
  readr::read_csv("http://tiny.cc/dcf/BibleNames.csv")
```
1. Which names have any of these words in them: “bar”, “dam”, “lory”?
```{r}
BibleNames %>%
  filter(grepl("bar|dam|lory",name))
```
2. Which names end with those words?
```{r}
BibleNames %>%
  filter(grepl("bar$|dam$|lory$",name))
```

### Problem 17.8

```{r}
CrimeSample <- 
  readr::read_csv("http://tiny.cc/dcf/Boston-Crimes-50.csv")
my_regex <- "(\\([+-]*[0-9]*[0-9\\.][0-9]*\\))[,](\\([+-]*[0-9]*[0-9\\.][0-9]*\\))"
CrimeSample %>%
  tidyr::extract("Location", into=c("lat", "long"), 
                 regex = my_regex, 
                 convert = TRUE)
```
