---
title: "STAT 184 Homework Exercises" 
subtitle: "DataComputing Chapters 10"
author: "Haichen Wei"
output: html_notebook
---

## Assigned Exercises

- DC Chapter 10 Exercises: 10.1, 10.2, 10.3, 10.4, 10.5, 10.6, 10.7  

## Front Matter

```{r}
# clean up workspace environment
rm(list = ls())

# all packages used for the assignment
library(mosaic)
library(tidyverse)

```



## Chapter 10 Exercises

### Problem 10.1

1. Summary Function
2. Data Verb
3. Transformation
4. Reduction Function
5. Data Verb
6. Summary Function
7. Data Verb
8. Summary Function

### Problem 10.2

1. mutate()
2. arrange()
3. select()
4. filter()
5. group by()
   summarise()

### Problem 10.3

1. nrow()
2. name()
3. help()
4. library()
5. group_by()

### Problem 10.4

1. BabyNames %>% filter(sex=="F") %>% arrange(count) 
2. BabyNames %>% filter(year > 1990, sex=="F")
3. BabyNames %>% filter(year %in% c(2012), sex=="M")
4. BabyNames %>% summarise(total=n())
5. BabyNames %>% select(name, count)

### Problem 10.5

1. 
```{r}
library(DataComputing)
data("Minneapolis2013")
Minneapolis2013 %>%
  group_by(First, Second) %>%
  summarise(total = n())
```

2. 
```{r}
Minneapolis2013 %>%
  group_by(First, Second) %>%
  summarise(total = n()) %>%
  arrange(desc(total)) %>%
  head(3)
```

3. 
```{r}
Minneapolis2013 %>%
  filter(First == "OLE SAVIOR") %>%
  group_by(Second) %>%
  summarise(total = n()) %>%
  arrange(desc(total)) %>%
  head(3)
```

### Problem 10.6

1. Color J is the largest on average (in terms of carats).
```{r}
library(ggplot2)
data("diamonds")
help("diamonds")
diamonds %>% 
  group_by(color) %>% 
  summarise(avg = mean(carat)) %>%
  arrange(desc(avg))
  head(1)
```
2. VVS1 has the largest average “table” per carat.
```{r}
diamonds %>% 
  group_by(clarity) %>% 
  summarise(avg2 = mean(table/carat)) %>%
  arrange(desc(avg2))
  head(1)
```

### Problem 10.7

```{r}
Minneapolis2013 %>%
  group_by(First) %>%
  summarise(total = n()) %>%
  mutate(candidate = ifelse(total < 5000, "minor", "major"))
```
