---
title: "Street or Road?"
author: "Haichen Wei"
date: "Nov 19, 2020"
output: 
    html_notebook
---


# Set Up

### Clean up environment and load packages
```{r}
# clean up the RStudio environment 
rm(list = ls())

# load all packages here: `mosaic`, `tidyverse`, `lubridate`, and all others used
library(tidyverse)
library(mosaic)
library(DataComputing)
```


### Solved Example

```{r}
download.file(url="http://tiny.cc/dcf/CMS_ProvidersSimple.rds",
              destfile = "YourNameForTheFile.rds")
DataTable <- readRDS("YourNameForTheFile.rds")
Addresses <- read_csv("http://tiny.cc/dcf/street-addresses.csv")
Sample <- 
  Addresses %>%
  sample_n(size = 50)
Sample
```

```{r}
Matches <- 
  Sample %>%
  filter(grepl(pattern = "PO", address))
Matches
```

```{r}
Dont <- 
  Sample %>%
  filter( ! grepl(pattern = "PO", address))
Dont
```

```{r}
pattern <- "BOX\\s+(\\d+)"

Matches <- 
  Sample %>% 
  filter(grepl(pattern, address))
Matches

Dont <-
  Sample %>%
  filter( ! grepl(pattern, address))
Dont
```

```{r}
BoxNumbers <- 
  Sample %>%
  filter(grepl(pattern, address)) %>%
  tidyr::extract(address, into = "boxnum", regex = pattern)
BoxNumbers
```

# Back to the Streets

### explain each line of code in English

```{r}
# make a regular expression string named pattern, the string inside parentheses is pattern that we want to extract, the string means that a ST or RD or ROAD in an address
pattern <- "(ST|RD|ROAD)" 

#create a new data frame table named LeftOvers
LeftOvers <-
  Addresses %>% 
  filter( ! grepl(pattern, address), # Find cases that don’t contain a ST or RD or ROAD in an address
          ! grepl("\\sAPT|UNIT\\s[\\d]+$", address), # Find cases that don’t contain APT or UNIT (within the space in front and behind it) which followed by one or more digits in an address
          ! grepl(" BOX ", address) # Find cases that don’t contain the BOX(within the space in front and behind it) in an address
          )
LeftOvers
```

### expand to match several more patterns (at least 12 total)

```{r}
# 1. Find cases that contain LN or LANE in an address
LANE <-
  Addresses %>% 
  filter( grepl("LN|LANE", address)) 
LANE
# 2.Find cases that contain AVE or AVENUE in an address
AVENUE <-
  Addresses %>% 
          filter(grepl("AVE|AVENUE", address))
AVENUE
# 3. Find cases that contain the TRAIL or TRL in an address
TRAIL <-
  Addresses %>% 
  filter(grepl("TRAIL|TRL", address)) 
TRAIL
# 4. Find cases that contain COURT or CT in an address
COURT <-
  Addresses %>% 
  filter(grepl("COURT|CT", address))
COURT
# 5. Find cases that contain CIRCLE or CIR in an address
CIRCLE <-
  Addresses %>% 
  filter(grepl("CIRCLE|CIR", address))
CIRCLE
# 6. Find cases that contain DR or DRIVE in an address
DRIVE <-
  Addresses %>% 
  filter(grepl("DR|DRIVE", address))
DRIVE
# 7. Find cases that contain WAY in an address
WAY <-
  Addresses %>% 
  filter(grepl("WAY", address))
WAY
# 8. Find cases that contain PATH in an address
PATH <-
  Addresses %>% 
  filter(grepl("PATH", address))
PATH
# 9. Find cases that contain POINT or PT in an address
POINT <-
  Addresses %>% 
  filter(grepl("POINT|PT", address))
POINT
# 10. Find cases that contain BLVD or Boulevard in an address
Boulevard <-
  Addresses %>% 
  filter(grepl("BLVD|Boulevard", address))
Boulevard
# 11. Find cases that contain PLACE or PL in an address
PLACE <-
  Addresses %>% 
  filter(grepl("PLACE|PL", address))
PLACE
# 12. Find cases that contain PKWY or Parkway in an address
Parkway <-
  Addresses %>% 
  filter(grepl("PKWY|Parkway", address)) 
Parkway
```

### provide a table in descending order of popularity for the street name identifiers you found

```{r}
LANE_NUM <- 
  LANE %>% summarise(total=n()) %>% mutate(name = "LN") 

AVENUE_NUM <- 
  AVENUE %>% summarise(total=n()) %>% mutate(name = "AVE") 
  
TRAIL_NUM <- 
  TRAIL %>% summarise(total=n()) %>% mutate(name = "TRAIL")
  
CIRCLE_NUM <- 
  CIRCLE %>% summarise(total=n()) %>% mutate(name = "CIRCLE")
  
COURT_NUM <- 
  COURT %>% summarise(total=n()) %>% mutate(name = "COURT")
  
DRIVE_NUM <- 
  DRIVE %>% summarise(total=n()) %>% mutate(name = "DR") 
  
PATH_NUM <- 
  PATH %>% summarise(total=n()) %>% mutate(name = "PATH") 
  
WAY_NUM <- 
  WAY %>% summarise(total=n()) %>% mutate(name = "WAY") 
  
POINT_NUM <- 
  POINT %>% summarise(total=n()) %>% mutate(name = "POINT") 
  
Boulevard_NUM <- 
  Boulevard %>% summarise(total=n()) %>% mutate(name = "BLVD") 
  
PLACE_NUM <- 
  PLACE %>% summarise(total=n()) %>% mutate(name = "PL") 
 
Parkway_NUM <- 
  Parkway %>% summarise(total=n()) %>% mutate(name = "PKWY")
  
Streets <- rbind(LANE_NUM, AVENUE_NUM, TRAIL_NUM, CIRCLE_NUM, COURT_NUM, DRIVE_NUM, PATH_NUM, WAY_NUM, POINT_NUM, Boulevard_NUM, PLACE_NUM, Parkway_NUM)
PopStName <-
Streets %>% 
  arrange( desc(total) )
PopStName
```

### use ggplot to construct a bar chart in descending order of popularity for the street name identifiers you found.

```{r}
ggplot(data=PopStName,aes(x=reorder(name,total),y=total ,fill=name)) + geom_bar(stat='identity',position='stack', width=.9) 
```
